\documentclass[fleqn,12pt]{wlscirep}

%\newcommand{\charly}[1]{\textcolor{?}{#1}}
\definecolor{gainsboro}{rgb}{0.86, 0.86, 0.86}
\newcommand{\doing}[1]{\textcolor{gray}{#1}}
\newcommand{\todo}[1]{\textcolor{gainsboro}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\fig}[1]{\textcolor{black}{#1}}
\newcommand\note[1]{\mbox{}\marginpar{\footnotesize\raggedright\hspace{0pt}\color{orange}\emph{#1}}}

\usepackage{lineno}
\usepackage{graphicx}

\let\oldthebibliography\thebibliography
\let\endoldthebibliography\endthebibliography
\renewenvironment{thebibliography}[1]{
  \begin{oldthebibliography}{#1}
    \setlength{\itemsep}{0em}
    \setlength{\parskip}{0em}
}
{
  \end{oldthebibliography}
}


\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = black
}

\usepackage{parskip}
\setlength{\parskip}{6pt}
\setlength{\parindent}{0pt}
\renewcommand{\baselinestretch}{1.1} 

\usepackage{wrapfig}

\title{Quantifying plasmid dynamics using single-cell microfluidics and image bioinformatics}



\author[1]{JCR Hern\'andez}
\author[2]{J Rodr\'iguez-Beltr\'an}
\author[2]{A San Mill\'an}
\author[1,*]{R Pe\~na-Miller}
\author[1,*]{A Fuentes-Hern\'andez}
\affil[1]{  \, Laboratorio de Biolog\'ia Sint\'etica y de Sistemas, Centro de Ciencias Gen\'omicas, Universidad Nacional Aut\'onoma de M\'exico, 62210, Cuernavaca, M\'exico}
\affil[2]{ \, Department of Microbiology, Hospital Universitario Ramon y Cajal (IRYCIS), Madrid, Spain}
\affil[*]{Corresponding authors: ayarifh@ccg.unam.mx, rpm@ccg.unam.mx}


\keywords{plasmid dynamics, fluorescence microscopy, image processing, microfluidics}

\begin{abstract}
\footnotesize{
Multicopy plasmids play an important role in bacterial ecology and evolution by accelerating the rate of adaptation and providing a platform for rapid gene amplification and evolutionary rescue. 
Despite the relevance of plasmids in bacterial evolutionary dynamics, evaluating the population-level consequences of randomly segregating and replicating plasmids in individual cells remains a challenging problem, both in theory and experimentally.
In recent years, technological advances in fluorescence microscopy and microfluidics have allowed studying temporal changes in gene expression by quantifying the fluorescent intensity of individual cells under controlled environmental conditions.
In this paper, we will describe the manufacture, experimental setup, and data analysis pipeline of different microfluidic systems that can be used to study plasmid dynamics, both in single-cells and in populations.
To illustrate the benefits and limitations of microfluidics to study multicopy plasmid dynamics, we will use an experimental model system consisting on {\em Escherichia coli} K12 carrying non-conjugative, multicopy plasmids ($19$ copies per cell, in average) encoding different fluorescent markers and $\beta$-lactam resistance genes. 
First, we will use an image-based flow cytometer to estimate changes in the allele distribution of a heterogeneous population under different selection regimes.  Then we will use a mother-machine microfluidic device to obtain time-series of fluorescent intensity of individual cells to argue that plasmid segregation and replication dynamics are inherently stochastic processes.  Finally, using a microchemostat, we track thousands of cells in time to reconstruct bacterial lineages and evaluate the plasmid frequency distributions that emerge in response to a range of selective pressures.
}
\end{abstract}
\begin{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Research highlights}

\begin{itemize}
    \item  Imaging can be used to study plasmid dynamics at multiple scales: from single-cells to bacterial populations 
    
    \item Intracellular plasmid dynamics is a noise-driven process resulting from randomly segregating and replicating plasmids.
    
    \item Selection modulates allele distributions by modifying the fraction of cells with different plasmid configurations.
    
    \item  Heteroplasmy is unstable in constant environments and stabilized in fluctuating environments.
    
\end{itemize}

\end{document}